---
title: "Week2_assign2_MoviewReview"
author: "Jayanth and Sarat"
date: "9/7/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Working with Text file and DB

### Step1:
- Reviews were collected using the Survey Mokney website, but text file was created manually using the data from this site as file extraction requires a subscription.
- The Source file was formatted as a tab Delimited file with no header data. File "MovieReviews.txt" is available at "https://github.com/jayanth0572/R_practice_and_assignments/MovieReviews.txt"

### Step2:
- postgres was installed on Windows machine and Database was created.
- Table named "movie_reviews" was created under Public schema to import data from the file
- File was imported using Copy command
- SQL file is available at "https://github.com/jayanth0572/R_practice_and_assignments/movie_review.sql"

> Create table statement
`CREATE TABLE public.movie_reviews
(
    "UserName" character varying(50) COLLATE pg_catalog."default" NOT NULL,
    "Movie" character varying(50) COLLATE pg_catalog."default" NOT NULL,
    "StarRating" character varying(5) COLLATE pg_catalog."default",
    "Review" character varying(250) COLLATE pg_catalog."default",
    "ReviewDate" date NOT NULL
)`

> Statement to import data into DB table
`COPY public.movie_reviews FROM 'C:\Users\Public\Downloads\MovieReviews.txt';`



# Loading the data from DB into R Dataframe

### Step1:
####Install packages
- for Database Interface between R and PostgreSQL installing package ("RPostgreSQL")
- for data manipulation functinos installing package "dplyr". ALternatively we can install whole "tidyverse" package.
```{r}
##install.packages("RPostgreSQL")
##install.packages("dplyr")
```
 

### Step2:
#### Establishing connection with DB server using the "PostgreSQL" driver
- First Driver object is created for reuse. We can also skip this step and use the dbDriver function directly in the dbConnect function.
- Next Connection to DB is established by using "dbConnect" function. This requires database details as well as login credentials.
```{r include=FALSE}
pwd <- "Removedtheactualpassword"
```

```{r}

library("RPostgreSQL")

#create driver object
drv <- dbDriver("PostgreSQL")

#create connection object
MovieReview_db <- dbConnect(drv,
                 user="postgres", 
                 password=pwd,
                 host="localhost", 
                 port=5432, 
                 dbname="assignments")

# list all the tables in "MovieReview_db" object
dbListTables(MovieReview_db)

```


### Step3:
#### Load Data to Dataframe and perform some analysis
- "dbGetQuery" fucntion is used to get data from DB
- Empty values in StarRating column were replaced with 0
- Calculated the avg. movie rating for each of the movies. "magritter" library was used fo this purpose.
```{r}

library(dplyr)
library(magrittr)

# query the database and load to a dataframe
MovieReview_df <- dbGetQuery(MovieReview_db, "SELECT * from public.movie_reviews;")

#Display Daframe values
#MovieReview_df

#replacing the empty columns in StarRating column with 0 for processing purposes.
MovieReview_df$star_rating[MovieReview_df$star_rating == ""] <- "0"

#Calculating the Average Movie rating for each movie. Pipe operator was used to make use of GroupBy and summarise functions.
AvgOfMovieReview <- MovieReview_df %>% group_by(movie) %>% summarise(star_rating = mean(as.numeric(star_rating)))

#Display Result
AvgOfMovieReview

```


### Step4:
- Disconnect from DB
```{r}

# disconnect from database
dbDisconnect(MovieReview_db)

```


